Evolutionary rescue and reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease

### 2023-05-19 doc/manuscript/translocation_bioarxiv.tex
Address formatting comments on compiled manuscript

% Please add a significance statement to explain the relevance of your work
\significancestatement{Understanding how species persist despite accelerating global change is
critical for the conservation of biodiversity. One possibility is rapid
evolution that restores positive growth to declining populations
(``evolutionary rescue''). Individuals from these rescued populations
could also be used to reestablish extinct populations. We evaluated both
possibilities in a once-common frog species driven to the brink of
extinction by a disease that has devastated global amphibian
biodiversity and for which few options exist to reverse disease impacts.
We found that frogs have evolved increased resistance to infection,
facilitating population recovery. In addition, introduction of frogs
from these populations allowed reestablishment despite the ongoing
presence of disease. Such reintroductions could play an important role
in biodiversity conservation in our rapidly changing world.}

@@

One consequence of human alteration of the biosphere is an increase in
emerging infectious diseases \citep{jones2008, fisher2012}. Such
diseases pose a severe threat to wildlife populations
\citep{daszak2000}, and have caused dramatic declines and extinctions in
a wide range of taxa, including echinoderms, mammals, birds, and
amphibians \citep{hewson2014, samuel2015, scheele2019, cunningham2021}.
Amid this global loss of biodiversity, some populations decimated by
disease have subsequently recovered to varying extents
\citep{newell2013, voyles2018, knapp2016}. Although we know relatively
little about the mechanisms underlying recovery
\citep{brannelly2021, russell2020}, host evolution is one possibility
\citep{carlson2014, searle2020}. This process is known from a diversity
of wildlife populations
\citep{savage2016, epstein2016, gignoux-wolfsohn2021, holland2022}, and
often results in increased host resistance and/or tolerance (see
\citep{medzhitov2012} for definitions). However, linking these
evolutionary responses to population recoveries is difficult and few
clear examples exist. In addition to promoting recovery of
disease-impacted populations, the resistant/tolerant individuals in such
populations could potentially be used to reestablish populations that
were extirpated following disease emergence
\citep{mendelson2019, brannelly2021}. To date, the use of
resistant/tolerant individuals for species recovery is rare \citep[but
see][]{joseph2018}, and studies that evaluate its potential utility are
critically needed.

@@

Results from the individual variant and splined window analyses show
that recovering populations have signatures of strong selection on
multiple regions of the genome. The analysis of individual variants
identified 11 ``outlier'' SNPs (i.e., showing significantly different
allele frequencies between naive versus recovering populations) from 7
distinct genes across 4 contigs (Figure S1B,C). One of the 7 identified genes (LOC108802036) did not have an
associated annotation. For the outlier SNPs, frequency differences
between the naive and recovering populations ranged from 41\% to 86\%.
Most of these SNPs showed frequency differences in only a subset of the
sampled populations (Figure~\ref{fig-allelefrequencies} A, B), but the
SNP in the RIN3 gene showed consistent differences in frequencies across
all populations (Figure~\ref{fig-allelefrequencies} C). This is
suggestive of strong - and parallel - selection at this locus associated
with exposure of frogs to Bd across multiple populations. (See SI
Appendix for additional description of these results.)

The splined window analysis identified 33 outlier regions for
\(\pi_{diff}\) and 58 outlier regions for \emph{F\textsubscript{ST}}
(Figure~\ref{fig-spline-manhattan} A, B). Of these, 9 regions were
outliers for both metrics (``shared regions'') and 2 of these shared
regions also contained one or more outlier SNPs (described above). A
total of 33 annotated genes were found in the 9 shared regions (SI
Appendix: Datasets S3, S4). The largest \(\pi_{diff}\), indicative of
directional selection, occurred in a 163kb region on Contig19, 12.9Mb
upstream of the RIN3 outlier SNP (Figure~\ref{fig-spline-manhattan} C).
This region contains approximately 500 SNPs and one annotated gene
called ``interferon-induced very large GTPase 1-like'' (GVINP1).
Additionally, a shared outlier region on Contig1 contained two
complement factor genes (C6 and C7). Interestingly, this region had a
large negative \(\pi_{diff}\) value, consistent with balancing
selection. Finally, one shared outlier region on Contig8 contained one
outlier SNP (TCF19) and was within 360kb of another outlier SNP (VARS)
(Figure~\ref{fig-spline-manhattan} D, Figure S2).
This region (854kb from the beginning of the outlier window to the VARS
SNP) contained a total of 8 annotated genes. In \emph{Xenopus}, five of
these genes occur in the extended major histocompatibility complex (MHC)
Class I region (FLOT1, TUBB, MDC1, CCHCR1, TCF19) and three occur in the
extended MHC Class III region (HSP70, LSM2, VARS) \citep{ohta2006}.
Therefore, this region under selection is part of the extended MHC Class
I and III complex and shows synteny with other amphibian genomes.

@@

To determine whether frogs from recovering populations could be used to
reestablish extirpated populations, we conducted 24 reintroductions in
Yosemite National Park (2006--2020). Each of the reintroductions
involved collection of adult frogs from 1 of 3 recovering, Bd-positive
``donor'' populations and translocating them to 1 of 12 nearby recipient
sites (Figure S3) \citep{seddon2014}. The donor
populations included 2 of the 5 recovering populations in the frog
evolution study described above (Figure~\ref{fig-allelefrequencies} A:
population 1 and 4). Following translocation, adult survival and
recruitment of new adults were estimated from CMR surveys and counts of
tadpoles and juveniles were obtained from visual encounter surveys
(VES).

Estimates of frog survival during the 1-year period following
translocation indicate that survival was highly variable between
recipient sites, but relatively constant within recipient sites (for the
subset of sites that received multiple translocations;
Figure~\ref{fig-translocation-survival}). As such, the average survival
of frogs in the first translocation to a site was an important predictor
of survival of frogs in subsequent translocations to that site
(Figure S4). This suggests an important
effect of site characteristics and/or (site * frog) interactions on frog
survival. In addition, 1-year survival was higher for frogs translocated
later in the study period than earlier
(Table S1), which we suggest resulted
primarily from our improved ability to choose recipient sites with
higher habitat quality for \emph{R. sierrae}. Finally, Bd loads were
fairly consistent before versus after translocation and loads were
nearly always well below the level indicative of severe chytridiomycosis
and associated frog mortality (Figure S5, see
SI Appendix text for additional details)
\citep{joseph2018, vredenburg2010}.

Of the 12 translocated populations, 9 (0.75) showed evidence of
successful reproduction in subsequent years, as indicated by the
presence of tadpoles and/or juveniles. For these 9 populations, one or
both life stages were detected in nearly all survey-years following
translocation (proportion of survey-years: median = 0.9, range =
0.29--1). These same populations were also those in which recruitment of
new adults (i.e., progeny of translocated individuals) was detected. As
with early life stages, recruits were detected in the majority of
post-translocation survey-years (proportion of survey-years: median =
0.79, range = 0.12--1). In summary, survey results suggest that
translocations resulted in the establishment of reproducing MYL frog
populations at most recipient sites despite the ongoing presence of Bd.

To determine whether Bd load had a negative effect on adult survival, as
would be expected if frogs were highly susceptible to Bd infection, we
developed linear and multilevel models. The best model of 1-year frog
survival identified several important predictors, but Bd load at the
time of translocation was not among them (bd\_load;
Figure S6). Important predictors included
winter severity in the year following translocation (snow\_t1), site
elevation (elevation), and donor population (donor\_72996;
Figure~\ref{fig-cond-effects}). Males had somewhat higher survival than
females, but this effect was only marginally important (sex\_male;
Figure~\ref{fig-cond-effects}, Figure S6).
Collectively, these results are consistent with frogs in recovering
populations having evolved sufficient resistance/tolerance to minimize
the adverse effects of Bd infection. In addition, the moderate Bd loads
that characterized frogs following translocation
(Figure S5) and relatively high survival of
frogs in numerous translocated cohorts
(Figure~\ref{fig-translocation-survival}) suggests that frogs
translocated from recovering populations can maintain the benefits of
resistance/tolerance in non-natal habitats. Finally, in 3 locations
where longer CMR time series allowed us to assess the survival of new
adults recruited to the population, naturally recruited adults had
equivalent or higher survival probabilities than the originally
translocated adults (Figure S7). Therefore,
frog resistance/tolerance in non-natal habitats is maintained across
generations.

@@

Given observed yearly adult survival probabilities (from analysis of CMR
data) and a baseline yearly recruitment probability \(\omega\) of 0.3
(i.e., the probability that, following survival, juveniles successfully
recruit to the adult class), our stage-structured model predicted that
six of twelve translocated populations should experience a long-run
growth rate \(\lambda\) greater than 1 in the presence of Bd
(Figure~\ref{fig-viability} A; median predicted \(\lambda\) ranges from
1.15-1.36 for these six populations). These six populations all had
observed yearly adult survival greater than 0.5. When we reduced the
probability of successful recruitment \(\omega\) to 0.1 (a highly
conservative estimate), long-run growth rate of five of these six
populations was still greater than 1 (Figure~\ref{fig-viability} A).

Even when incorporating (i) demographic stochasticity and (ii)
environmental stochasticity in the probability of juvenile recruitment
to the adult class \(\omega\) (the transition that we expect to be the
most subject to environmental variability in the presence of Bd),
populations with high adult survival are likely to persist over a 50
year time horizon. Our model predicted that, following a single
introduction of 40 adult individuals into a population and using the
baseline value for \(\omega\) of 0.3, the six populations with the
highest adult survival probabilities had 50-year extinction
probabilities of 0 to 0.32 (Figure~\ref{fig-viability} B). This
indicates strong potential for long-term persistence in the presence of
Bd and environmental variability in recruitment. When we decreased
\(\omega\) from 0.3 to a conservative estimate of 0.15, 50-year
extinction probability remained less than 0.5 for three populations and
was 0.79 to 0.94 for three additional populations
(Figure~\ref{fig-viability} B). In contrast, for the six populations
where yearly adult survival probability \(\sigma_{A_R} < 0.5\),
extinction probability over 50 years was always predicted to be 1. To
test the validity of our model predictions, we demonstrated that our
stochastic model could describe the general recovery trajectory of our
translocated population with the longest survey history
(Figure~\ref{fig-viability} C; population 70550, surveyed for 16 years).

@@

Disease-induced population declines are decimating global biodiversity
\citep{daszak2000}, but strategies to recover affected species are
generally lacking. Here, we tested the possibility that populations of
resistant/tolerant individuals could be used to reestablish extirpated
populations. Although previously proposed
\citep{mendelson2019, scheele2021}, this strategy has rarely been used
in the wild (see \citep{joseph2018} for a small-scale example). In
particular, although numerous studies have documented signatures of
selection in hosts following disease-induced declines
\citep[e.g.,][]{epstein2016, gignoux-wolfsohn2021}, we are unaware of
any that used individuals bearing the resulting adaptive alleles to
reestablish extirpated populations. The results of our study provide
evidence that (1) MYL frogs have experienced strong selection following
exposure to a recently-emerged lethal fungal disease and that some of
the genomic regions under selection are associated with genes related to
disease resistance/tolerance, and (2) reintroduction of individuals from
these resistant/tolerant populations allows the successful
reestablishment of viable frog populations despite ongoing disease.

Previous field studies in MYL frogs show that frog-Bd dynamics and frog
survival in the presence of Bd are fundamentally different between naive
and recovering populations. Following the arrival and establishment of
Bd in previously-naive populations, adult frogs develop high Bd loads
that lead to mass die-offs \citep{vredenburg2010}. In contrast, in
recovering populations adult frogs typically have low-to-moderate and
relatively constant Bd loads and mass die-offs are not observed
\citep[see also
Figure S5]{briggs2010, knapp2011}. The
differences in Bd load of frogs from naive and recovering populations
are also observed in controlled laboratory studies \citep[Figure
4]{knapp2016}. The observed reduced susceptibility of frogs from
recovering populations could in theory be due to several factors,
including natural selection for more resistant/tolerant genotypes and/or
acquired immunity, but until now evidence to evaluate the role of
evolution was lacking.

Results from our genomic analyses strongly suggest that natural
selection for adaptive alleles and resulting increased
resistance/tolerance is at least partially responsible for the reduced
susceptibility of frogs in recovering populations. We identified
specific alleles and genomic regions showing signatures of selection
between adjacent Bd-naive and recovering MYL frog populations,
consistent with strong selection following Bd exposure. Importantly,
some regions under selection are associated with cellular and
immunological mechanisms known to contribute to disease resistance,
including in amphibians \citep{zamudio2020}. For example, the MHC plays
an important role in immunity. In our study, we identified a region that
shows evidence of selection in recovering populations and contains eight
genes associated with either the MHC Class I or Class III regions. These
results corroborate numerous previous studies linking MHC genes to
amphibian resistance/tolerance against Bd
\citep[e.g.,][]{savage2011, bataille2015}. Similarly, the region with
the strongest indication of directional selection (as measured by
\(\pi_{diff}\)) contains the interferon-related gene GVINP1. Several
previous studies have found this gene to be differentially expressed
during Bd infection \citep[e.g.,][]{grogan2018, ellison2014} and in
populations differing in Bd susceptibility \citep{grogan2018}. This gene
is also strongly linked to disease in salmon, explaining a notable 20\%
of the resistance phenotype \citep{robledo2020, robledo2018}. We also
identified a region, characterized by high \emph{F\textsubscript{ST}}
and low \(\pi_{diff}\), that contained the complement genes C6 and C7.
This could indicate that balancing selection is acting in this region of
the genome to favor a diverse set of alleles, as is known for C6 in
humans \citep{soejima2005}. Finally, based on the analysis of individual
outlier variants, the RIN3 gene showed a consistent pattern of allele
frequency differences across all nine of the frog populations sampled in
this study, indicating consistent selection in populations distributed
across a wide geographic area. This gene is associated with inflammation
response and in \emph{Xenopus} is highly expressed in wound stumps
following amputation of tadpole tails \citep{fukazawa2009}. Also in this
analysis, the outlier variant with the lowest p-value (uncharacterized
gene LOC108802036) is located adjacent to a type I interferon gene
(Np-IFNi2) in another frog genome \citep{gan2018}, and together with
GVINP1 further indicates the importance of interferon-related genes in
this system.

@@

Despite the resistance/tolerance of adult MYL frogs in recovering
populations, individual and population-level impacts of Bd are still
evident. In an earlier study of 2 of our 12 translocated populations
\citep{joseph2018}, Bd infection and load had detectable effects on the
survival of adults and may have influenced population establishment
(sites referred to as ``Alpine'' and ``Subalpine'' in \citep{joseph2018}
are identified as ``70550'' and ``70505'' in the current study).
Applying similar analyses to all 12 of our translocated populations
would likely provide a broader perspective of the ongoing effect of Bd.
In addition to these important but relatively subtle effects of Bd on
adults, the impacts on younger life stages are more apparent. MYL frogs
immediately following metamorphosis (``metamorphs'') are highly
susceptible to Bd infection \citep{ellison2018} and as a result
experience high mortality. This high susceptibility of metamorphs is
documented for numerous species of anurans, and may result from the
poorly developed immune system characteristic of this life stage
\citep{humphries2022}. In naturally recovering and translocated MYL frog
populations, we suggest that the high mortality of metamorphs is an
important limitation on subsequent recruitment of new adults. Therefore,
although adult MYL frogs appear relatively resistant/tolerant, Bd
infection continues to have important limiting effects on recovering
populations \citep[see also][]{hollanders2022}.

@@

We collected DNA samples via buccal swabbing \citep{broquet2007} from 53
\emph{Rana muscosa}/\emph{Rana sierrae} individuals: 24 from four naive
populations, and 29 from five recovering populations. These populations
were located in the southern Sierra Nevada, from northern Yosemite
National Park to northern Sequoia National Park
(Figure~\ref{fig-allelefrequencies}). Samples were collected from 5-6
frogs per population. To minimize potential confounding effects caused
by variation in frog genotypes across latitude
\citep[e.g.,][]{rothstein2020}, we selected sampling sites such that
both population types were represented across similar latitudinal
ranges. DNA was extracted following Qiagen DNEasy manufacturer's
protocols. We sequenced the samples using an exome capture assay as
described in Byrne et al.~(in review). Briefly, genomic libraries were
prepared and captured using a custom Nimblegen capture pool. Capture
baits were designed based on the coding regions of the \emph{R. muscosa}
transcriptome (GenBank accession GKCT00000000). Captured libraries were
then pooled and sequenced on a NovaSeq 6000 150PE Flow Cell S1 at the
Vincent J. Coates Genomics Sequencing Lab at UC Berkeley.

@@

To visualize the genomic relationship of our populations we conducted a
PCA using the glPCA function in the adegenet R package
\citep{jombart2008}. To detect regions of the genome under selection
between naive and recovering populations, we used two approaches: (1) a
multivariate linear mixed model to evaluate individual variants (SNPs
and INDELs), and (2) a splined window analysis to evaluate larger
genomic regions. For the variant analysis, we first used a stringent
data filter to include only variants with \textless{} 5\% missing data
(missing for no more than 2 individuals), and then calculated the
likelihood ratio statistic for the resulting set of 148,307 high quality
variants across 127 contigs using GEMMA \citep{zhou2014}. GEMMA
calculates and incorporates a relatedness matrix for input samples,
allowing us to account for relatedness and population structure when
calculating likelihood ratio statistics. We identified variants showing
different allele frequencies between naive versus recovering populations
(``outliers'') using a Bonferroni- corrected significance level of 0.01.
We visualized the results using a Manhattan plot and qqplot. We
developed a more liberal set of outlier variants using a Bonferroni-
corrected significance level of 0.05 and used this set solely for the
gene ontology (GO) analysis (see below and SI Appendix: Dataset S1, S2).

>>>>

% Please add a significance statement to explain the relevance of your work
% \significancestatement{Understanding how species persist despite accelerating global change is
% critical for the conservation of biodiversity. One possibility is rapid
% evolution that restores positive growth to declining populations
% (``evolutionary rescue''). Individuals from these rescued populations
% could also be used to reestablish extinct populations. We evaluated both
% possibilities in a once-common frog species driven to the brink of
% extinction by a disease that has devastated global amphibian
% biodiversity and for which few options exist to reverse disease impacts.
% We found that frogs have evolved increased resistance to infection,
% facilitating population recovery. In addition, introduction of frogs
% from these populations allowed reestablishment despite the ongoing
% presence of disease. Such reintroductions could play an important role
% in biodiversity conservation in our rapidly changing world.}

@@

One consequence of human alteration of the biosphere is an increase in
emerging infectious diseases \citep{jones2008, fisher2012}. Such
diseases pose a severe threat to wildlife populations
\citep{daszak2000}, and have caused dramatic declines and extinctions in
a wide range of taxa, including echinoderms, mammals, birds, and
amphibians \citep{hewson2014, samuel2015, scheele2019, cunningham2021}.
Amid this global loss of biodiversity, some populations decimated by
disease have subsequently recovered to varying extents
\citep{scheele2017, voyles2018, knapp2016}. Although we know relatively
little about the mechanisms underlying recovery
\citep{brannelly2021, russell2020}, host evolution is one possibility
\citep{carlson2014, searle2020}. This process is known from a diversity
of wildlife populations
\citep{savage2016, epstein2016, gignoux-wolfsohn2021, holland2022}, and
often results in increased host resistance and/or tolerance \citep[see][
for definitions]{medzhitov2012}. However, linking these evolutionary
responses to population recoveries is difficult and few clear examples
exist. In addition to promoting recovery of disease-impacted
populations, the resistant/tolerant individuals in such populations
could potentially be used to reestablish populations that were
extirpated following disease emergence
\citep{mendelson2019, brannelly2021}. To date, the use of
resistant/tolerant individuals for species recovery is rare \citep[but
see][]{joseph2018}, and studies that evaluate its potential utility are
critically needed.

@@

Results from the individual variant and splined window analyses show
that recovering populations have signatures of strong selection on
multiple regions of the genome. The analysis of individual variants
identified 11 ``outlier'' SNPs (i.e., showing significantly different
allele frequencies between naive versus recovering populations) from 7
distinct genes across 4 contigs (Figure S1B, C;
SI Appendix: Datasets S1, S2). One of the 7 identified genes
(LOC108802036) did not have an associated annotation. For the outlier
SNPs, frequency differences between the naive and recovering populations
ranged from 41\% to 86\%. Most of these SNPs showed frequency
differences in only a subset of the sampled populations
(Figure~\ref{fig-allelefrequencies}A, B), but the SNP in the RIN3 gene
showed consistent differences in frequencies across all populations
(Figure~\ref{fig-allelefrequencies}C). This is suggestive of strong -
and parallel - selection at this locus associated with exposure of frogs
to Bd across multiple populations.

The splined window analysis identified 33 outlier regions for
\(\pi_{diff}\) and 58 outlier regions for \emph{F\textsubscript{ST}}
(Figure~\ref{fig-spline-manhattan}A, B). Of these, 9 regions were
outliers for both metrics (``shared regions'') and 2 of these shared
regions also contained one or more of the outlier SNPs described above.
A total of 33 annotated genes were found in the 9 shared regions (SI
Appendix: Datasets S3, S4). Given this large number of genes, here we
focus on those with the strongest signal of selection and/or
immune-related functions. The largest \(\pi_{diff}\), indicative of
directional selection, occurred in a 163kb region on Contig19, 12.9Mb
upstream of the RIN3 outlier SNP (Figure~\ref{fig-spline-manhattan}C).
This region contains approximately 500 SNPs and one annotated gene
called ``interferon-induced very large GTPase 1-like'' (GVINP1).
Additionally, a shared outlier region on Contig1 contained two
complement factor genes (C6 and C7). Interestingly, this region had a
large negative \(\pi_{diff}\) value, consistent with balancing
selection. Finally, one shared outlier region on Contig8 contained one
outlier SNP (TCF19) and was within 360kb of another outlier SNP (VARS)
(Figure~\ref{fig-spline-manhattan}D, Figure S2).
This region (854kb from the beginning of the outlier window to the VARS
SNP) contained a total of 8 annotated genes. In \emph{Xenopus}, five of
these genes occur in the extended major histocompatibility complex (MHC)
Class I region (FLOT1, TUBB, MDC1, CCHCR1, TCF19) and three occur in the
extended MHC Class III region (HSP70, LSM2, VARS) \citep{ohta2006}.
Therefore, this region under selection is part of the extended MHC Class
I and III complex and shows synteny with other amphibian genomes.

@@

To determine whether frogs from recovering populations could be used to
reestablish extirpated populations, we conducted 24 reintroductions in
Yosemite National Park (2006--2020). Each of the reintroductions
involved collection of adult frogs from 1 of 3 recovering, Bd-positive
``donor'' populations and translocating them \citep{seddon2014} to 1 of
12 nearby recipient sites (Figure S3). The donor
populations included 2 of the 5 recovering populations in the frog
evolution study described above (Figure~\ref{fig-allelefrequencies}A:
population 1 and 4). Following translocation, adult survival and
recruitment of new adults was estimated from CMR surveys and counts of
tadpoles and juveniles were obtained from visual encounter surveys
(VES).

Estimates of frog survival during the 1-year period following
translocation indicate that survival was highly variable between
recipient sites, but relatively constant within recipient sites (for the
subset of sites that received multiple translocations;
Figure~\ref{fig-translocation-survival}). As such, the average survival
of frogs in the first translocation to a site was an important predictor
of survival of frogs in subsequent translocations to that site
(Figure S4). This suggests an important
effect of site characteristics and/or (site * frog) interactions on frog
survival. In addition, 1-year survival was higher for frogs translocated
later in the study period than earlier
(Table S1), which we suggest resulted
primarily from our improved ability to choose recipient sites with
higher habitat quality for \emph{R. sierrae}. Finally, Bd loads were
fairly consistent before versus after translocation and loads were
nearly always well below the level indicative of severe chytridiomycosis
and associated frog mortality (Figure S5)
\citep{joseph2018, vredenburg2010}. Additional details on methods used
to estimate Bd loads are provided in the SI Appendix.

Of the 12 translocated populations, 9 (0.75) showed evidence of
successful reproduction in subsequent years, as indicated by the
presence of tadpoles and/or juveniles. For these 9 populations, one or
both life stages were detected in nearly all survey-years following
translocation (proportion of survey-years: median = 0.9, range =
0.29--1). These same populations were also those in which recruitment of
new adults (i.e., progeny of translocated individuals) was detected. As
with early life stages, recruits were detected in the majority of
post-translocation survey-years (proportion of survey-years: median =
0.79, range = 0.12--1). In summary, survey results indicate that
translocations resulted in the establishment of reproducing MYL frog
populations at most recipient sites despite the ongoing presence of Bd.

To determine whether Bd load had a negative effect on adult survival, as
would be expected if frogs were highly susceptible to Bd infection, we
developed linear and multilevel models. The best model of 1-year frog
survival identified several important predictors, but Bd load at the
time of translocation was not among them
(Figure S6). Important predictors included
winter severity in the year following translocation, site elevation, and
donor population (Figure~\ref{fig-cond-effects}). Males had somewhat
higher survival than females, but this effect was only marginally
important (Figure~\ref{fig-cond-effects},
Figure S6). Collectively, these results are
consistent with frogs in recovering populations having evolved
sufficient resistance/tolerance to minimize the adverse effects of Bd
infection. In addition, the moderate Bd loads that characterized frogs
following translocation (Figure S5) and
relatively high survival of frogs in numerous translocated cohorts
(Figure~\ref{fig-translocation-survival}) suggests that frogs
translocated from recovering populations can maintain the benefits of
resistance/tolerance in non-natal habitats. Finally, in 3 locations
where longer CMR time series allowed us to assess the survival of new
adults recruited to the population, naturally recruited adults had
equivalent or higher survival probabilities than the originally
translocated adults (Figure S7). Therefore,
frog resistance/tolerance in non-natal habitats is maintained across
generations.

@@

Given observed yearly adult survival probabilities (from analysis of CMR
data) and a baseline yearly recruitment probability \(\omega\) of 0.3
(i.e., the probability that, following survival, juveniles successfully
recruit to the adult class), our stage-structured model predicted that
six of twelve translocated populations should experience a long-run
growth rate \(\lambda\) greater than 1 in the presence of Bd
(Figure~\ref{fig-viability}A; median predicted \(\lambda\) ranges from
1.15-1.36 for these six populations). These six populations all had
observed yearly adult survival greater than 0.5. When we reduced the
probability of successful recruitment \(\omega\) to 0.1 (a highly
conservative estimate), long-run growth rate of five of these six
populations was still greater than 1 (Figure~\ref{fig-viability}A).

Even when incorporating (i) demographic stochasticity and (ii)
environmental stochasticity in the probability of juvenile recruitment
to the adult class \(\omega\) (the transition that we expect to be the
most subject to environmental variability in the presence of Bd),
populations with high adult survival are likely to persist over a 50
year time horizon. Our model predicted that, following a single
introduction of 40 adult individuals into a population and using the
baseline value for \(\omega\) of 0.3, the six populations with the
highest adult survival probabilities (\(\sigma_{A_R} > 0.5\)) had
50-year extinction probabilities of 0 to 0.32
(Figure~\ref{fig-viability}B). This indicates strong potential for
long-term persistence in the presence of Bd and environmental
variability in recruitment. When we decreased \(\omega\) from 0.3 to a
conservative estimate of 0.15, 50-year extinction probability remained
less than 0.5 for three populations and was 0.79 to 0.94 for three
additional populations (Figure~\ref{fig-viability}B). In contrast, for
the six populations where yearly adult survival probability
\(\sigma_{A_R} < 0.5\), extinction probability over 50 years was always
predicted to be 1. To test the validity of our model predictions, we
demonstrated that our stochastic model could describe the general
recovery trajectory of our translocated population with the longest
survey history (Figure~\ref{fig-viability}C; population 70550, surveyed
for 16 years).

@@

Disease-induced population declines are decimating global biodiversity
\citep{daszak2000}, but strategies to recover affected species are
generally lacking. Here, we tested the possibility that populations of
resistant/tolerant individuals could be used to reestablish extirpated
populations. Although previously proposed
\citep{mendelson2019, scheele2021}, this strategy has rarely been used
in the wild \citep[see][ for a small-scale example]{joseph2018}. In
particular, although numerous studies have documented signatures of
selection in hosts following disease-induced declines
\citep[e.g.,][]{epstein2016, gignoux-wolfsohn2021}, we are unaware of
any that used individuals bearing the resulting adaptive alleles to
reestablish extirpated populations. The results of our study provide
evidence that (1) MYL frogs have experienced strong selection following
exposure to a recently-emerged lethal fungal disease and that some of
the genomic regions under selection contain genes related to disease
resistance/tolerance, and (2) reintroduction of individuals from these
resistant/tolerant populations allows the successful reestablishment of
viable frog populations despite ongoing disease.

Previous field studies in MYL frogs show that frog-Bd dynamics and frog
survival in the presence of Bd are fundamentally different between naive
and recovering populations. Following the arrival and establishment of
Bd in previously-naive populations, adult frogs develop high Bd loads
that lead to mass die-offs \citep{vredenburg2010}. In contrast, in
recovering populations adult frogs typically have low-to-moderate and
relatively constant Bd loads and mass die-offs are not observed
\citep[see also
Figure S5]{briggs2010, knapp2011}. The
differences in Bd load of frogs from naive and recovering populations
are also observed in controlled laboratory studies \citep[see Figure 4
in][]{knapp2016}. The observed reduced susceptibility of frogs from
recovering populations could in theory be due to several factors,
including natural selection for more resistant/tolerant genotypes and/or
acquired immunity, but until now evidence to evaluate the role of
evolution was lacking.

Results from our genomic analyses strongly suggest that natural
selection for adaptive alleles and resulting increased
resistance/tolerance is at least partially responsible for the reduced
susceptibility of frogs in recovering populations. We identified
specific alleles and genomic regions showing signatures of selection
between adjacent Bd-naive and recovering MYL frog populations,
consistent with strong selection following Bd exposure. Importantly,
some regions under selection are associated with cellular and
immunological mechanisms known to contribute to disease resistance,
including in amphibians \citep{zamudio2020}. For example, the MHC plays
an important role in immunity. In our study, we identified a region that
shows evidence of selection in recovering populations and contains eight
genes associated with either the MHC Class I or Class III regions. These
results corroborate numerous previous studies linking MHC genes to
amphibian resistance/tolerance against Bd
\citep[e.g.,][]{savage2011, bataille2015}. Similarly, the region with
the strongest indication of directional selection (as measured by
\(\pi_{diff}\)) contains the interferon-related gene GVINP1. Several
previous studies have found this gene to be differentially expressed
during Bd infection \citep[e.g.,][]{grogan2018, ellison2014} and in
populations differing in Bd susceptibility \citep{grogan2018}. This gene
is also strongly linked to disease in salmon, explaining a notable 20\%
of the resistance phenotype \citep{robledo2020, robledo2018}. We also
identified a region, characterized by high \emph{F\textsubscript{ST}}
and low \(\pi_{diff}\), that contained the complement genes C6 and C7.
The complement system plays an important role in innate immunity
\citep{rieraromo2016}, and our results could indicate that balancing
selection is acting in this region of the genome to favor a diverse set
of alleles, as is known for C6 in humans \citep{soejima2005}. Finally,
based on the analysis of individual outlier variants, the RIN3 gene
showed a consistent pattern of allele frequency differences across all
nine of the frog populations sampled in this study, indicating
consistent selection in populations distributed across a wide geographic
area. This gene is associated with inflammation response and in
\emph{Xenopus} is highly expressed in wound stumps following amputation
of tadpole tails \citep{fukazawa2009}. Also in this analysis, the
outlier variant with the lowest p-value (uncharacterized gene
LOC108802036) is located adjacent to a type I interferon gene (Np-IFNi2)
in another frog genome \citep{gan2018}, and together with GVINP1 further
suggests the importance of interferon-related genes in this system.

@@

Despite the resistance/tolerance of adult MYL frogs in recovering
populations, individual and population-level impacts of Bd are still
evident. In an earlier study of 2 of our 12 translocated populations
\citep{joseph2018}, Bd infection and load had detectable effects on the
survival of adults and may have influenced population establishment
(sites referred to as ``Alpine'' and ``Subalpine'' in \citep{joseph2018}
are identified as ``70550'' and ``70505'' in the current study).
Applying similar analyses to all 12 of our translocated populations
would likely provide a broader perspective of the ongoing effect of Bd.
In addition to these important but relatively subtle effects of Bd on
adults, the impacts on younger life stages are more apparent. MYL frogs
immediately following metamorphosis (``metamorphs'') are highly
susceptible to Bd infection \citep{ellison2018} and as a result
experience high mortality. This high susceptibility of metamorphs is
documented in numerous species of anurans, and may result from the
poorly developed immune system characteristic of this life stage
\citep{humphries2022}. In naturally recovering and translocated MYL frog
populations, we suggest that the high mortality of metamorphs is an
important limitation on subsequent recruitment of new adults. Therefore,
although adult MYL frogs appear relatively resistant/tolerant, Bd
infection continues to have important limiting effects on recovering
populations \citep[see also][]{hollanders2022}.

@@

We collected DNA samples via buccal swabbing \citep{broquet2007} from 53
\emph{Rana muscosa}/\emph{Rana sierrae} individuals: 24 from 4 naive
populations, and 29 from 5 recovering populations. These populations are
located in the southern Sierra Nevada, from northern Yosemite National
Park to northern Sequoia National Park
(Figure~\ref{fig-allelefrequencies}). Samples were collected from 5-6
frogs per population. To minimize potential confounding effects caused
by variation in frog genotypes across latitude
\citep[e.g.,][]{rothstein2020}, we selected sampling sites such that
both population types were represented across similar latitudinal
ranges. DNA was extracted following Qiagen DNEasy manufacturer's
protocols. We sequenced the samples using an exome capture assay as
described in Byrne et al.~(in review). Briefly, genomic libraries were
prepared and captured using a custom Nimblegen capture pool. Capture
baits were designed based on the coding regions of the \emph{R. muscosa}
transcriptome (GenBank accession GKCT00000000). Captured libraries were
then pooled and sequenced on a NovaSeq 6000 150PE Flow Cell S1 at the
Vincent J. Coates Genomics Sequencing Lab at UC Berkeley. Raw sequencing
reads are available from NCBI SRA (PRJNA870451).

@@

To visualize the genomic relationships of our populations we conducted a
PCA using the glPCA function in the adegenet R package
\citep{jombart2008}. To detect regions of the genome under selection
between naive and recovering populations, we used two approaches: (1) a
multivariate linear mixed model to evaluate individual variants (SNPs
and INDELs), and (2) a splined window analysis to evaluate larger
genomic regions. For the variant analysis, we first used a stringent
data filter to include only variants with \textless{} 5\% missing data
(missing for no more than 2 individuals), and then calculated the
likelihood ratio statistic for the resulting set of 148,307 high quality
variants across 127 contigs using GEMMA \citep{zhou2014}. GEMMA
calculates and incorporates a relatedness matrix for input samples,
allowing us to account for relatedness and population structure when
calculating likelihood ratio statistics. We identified variants showing
different allele frequencies between naive versus recovering populations
(``outliers'') using a Bonferroni-corrected significance level of 0.01.
We visualized the results using a Manhattan plot and qqplot. We
developed a more liberal set of outlier variants using a
Bonferroni-corrected significance level of 0.05 and used this set solely
for the gene ontology (GO) analysis (see below and SI Appendix: Dataset
S1, S2).
